---
title: "scraps"
output: html_document
---



# Notes to self

## Associations needed to hypoth-test

- $P(s_1)$: gender $\sim t_{si} \times \{l, c\} \times \{v, g\}$, with quartiles
- $P(s_2)$: gender $\sim t_{ic} \times \{l, c\} \times \{v, g\}$, with quartiles
- $\rho(a_1)$: Age $\sim t_{si} \times \{l, c\} \times \{v, g\}$
- $\rho(a_2)$: Age $\sim t_{ic} \times \{l, c\} \times \{v, g\}$
- $\rho(t): t_{si} \sim t_{ic} \times \{l, c\} \times \{v, g\}$

Or in notation,

$$ \{s_1, s_2, a_1, a_2, t\} \times \{l, c\} \times  \{v, g\} = |\mathbb{V}| \cdot |\mathbb{C}| \cdot |\mathbb{S}| = 20$$

where $\mathbb{V}$ is what pair of variables you're correlating, $\mathbb{C}$ is `cancer_type`, and $\mathbb{S}$ is `site`.

Maybe a notation better illustrating the pairing would be something like:

$$ \{st, su, at, au, tu\} \times \{l, c\} \times  \{v, g\}$$

var  |  sex  |   age   |  $t_{si}$   |   $t_{ic}$
-----|-------|---------|-------------|-------------
sex  |  x    |         |      s1     |     s2
age  |  x    |    x    |      a1     |     a2         
tsi  |  x    |    x    |      x      |     t         
tic  |  x    |    x    |      x      |     x




## Shell table for me

Too abstract for paper

```{r shell-table}
tribble(
  ~v1, ~v2, ~cancer, ~site,  ~pval,
  "Sex",  "t_si", "Lung", "Geisinger", "tk",
  "sex", "t_si", "Colon", "Geisinger",  "tk",
) %>% 
  kable()
```

It's really only 2 gender correlations ($s_1, s_2$) that we test with Wilcoxon, but it's times 4 strata, so that's 8 tests. Best to have a Wilcoxon function for this table.






# FIXME - Digression: Exploratory analysis

## GGally

```{r ggally-LONG, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
library(GGally)

reviewed_all %>% 
  filter(cancer_type=="Lung", site=="VA") %>% 
  select(MOD_definite, signal_to_init, init_to_completion) %>% 
  ggpairs() +
  labs(title="VA Lung")

reviewed_all %>% 
  filter(cancer_type=="Colon", site=="VA") %>% 
  select(MOD_definite, signal_to_init, init_to_completion) %>% 
  ggpairs() +
  labs(title="VA Colon")

reviewed_all %>% 
  filter(cancer_type=="Lung", site=="Geisinger") %>% 
  select(Age, Sex, MOD_definite, signal_to_init, init_to_completion) %>% 
  ggpairs() +
  labs(title="Geisinger Lung")

reviewed_all %>% 
  filter(cancer_type=="Colon", site=="Geisinger") %>% 
  select(Age, Sex, MOD_definite, signal_to_init, init_to_completion) %>% 
  ggpairs() +
  labs(title="Geisinger Colon")
```


## Sanity check

*Lesson:* Something weird about how cor.test handles NAs.

*Lesson:* Basically, there are (only) $n_l = 65$ Geisinger lung and $n_c = 70$ Geisinger colon with "temporality" data. $n = n_l + n_c = 135$. And the following predicates map 1:1 in the crosstabulation:

- `signal_present = "Yes"`
- `!is.na(sig_to_init)`
- `!is.na(init_to_complete)`

**Note to self!** Maybe (probably) I need to select only for those with red flag, meaning `signal_present = "Yes"`.

## How can you get negative numbers?

Workup completed (exactly one year) *before* workup initiated?

```{r negatives, eval=FALSE}
temporal_gv %>%  # FIXME
  filter(signal_present=="Yes", sig_to_init < 0 | init_to_complete < 0) %>%
  select(cancer_type, sig_to_init, init_to_complete, signal_date, dx_eval_init_date, dx_eval_complete_date) %>% 
  kable()
```

I suspect there are entry errors in year here.

## Correct stats: Non-NA `cor.test`

(Starting with correlation of `Age` and `sig_to_init`, like we did under "Sanity Check," above.)

```{r non-na-spearman, eval=FALSE}
t_all <- temporal_gv %>%  # FIXME
  filter(signal_present=="Yes", !is.na(sig_to_init)) %>%
  select(cancer_type, site, Age, sig_to_init, init_to_complete, Reviewer)

cor_ic <- cor.test(t_all$sig_to_init, t_all$init_to_complete, method = "spearman")
cor_ac <- cor.test(t_all$Age, t_all$init_to_complete, method = "spearman")
cor_ai <- cor.test(t_all$Age, t_all$sig_to_init, method = "spearman")

cor_ic
cor_ac
cor_ai

tribble(
  ~var1, ~var2, ~p.value, ~rho,
  "sig_to_init", "init_to_complete", cor_ic$p.value, cor_ic$estimate,
  "Age", "init_to_complete", cor_ac$p.value, cor_ac$estimate,
  "Age", "sig_to_init", cor_ai$p.value, cor_ai$estimate
) %>% 
  mutate(
    neg_log_p = round(-1 * log10(p.value), 1),
    across(c(p.value, rho), ~ round(., 3))
  ) %>% 
  kable()
```

## Interval plot

```{r wacky-time-plot, eval=FALSE}
t_all %>% 
  filter(cancer_type == "Lung") %>% 
  # arrange(sig_to_init + init_to_complete) %>% 
  arrange(sig_to_init) %>% 
  mutate(
    sig_to_complete = sig_to_init + init_to_complete,
    patient_id = row_number(),
    vmin = sig_to_init,
    vmax = sig_to_complete
  ) %>% 
  pivot_longer(
    cols = c("sig_to_init", "sig_to_complete"),
    names_to = "interval_type",
    names_prefix = "sig_to_"
  ) %>% 
  ggplot(aes(x = value, y = patient_id)) +
  geom_point(aes(shape = factor(interval_type), colour = factor(Reviewer))) +
  geom_linerange(aes(xmin = vmin, xmax=vmax, color=factor(Reviewer))) +
  geom_linerange(aes(xmin = 0, xmax=vmin, color=factor(Reviewer)), alpha = 0.2) +
  scale_x_sqrt()
```

## Reviewer bias?

```{r exploratory-reviewer-bias, eval=FALSE}
t_all %>% 
  pivot_longer(cols = c(sig_to_init, init_to_complete), names_to = "varname") %>% 
  filter(cancer_type=="Lung") %>% 
  group_by(Reviewer, varname) %>% 
  summarise(
    mean = mean(value),
    sd = sd(value),
    median = median(value),
    q1 = quantile(value, 0.25),
    q3 = quantile(value, 0.75)
  ) %>% 
  arrange(desc(varname)) %>% 
  kable()

r_si <- t_all %>% filter(cancer_type=="Lung",
                         Reviewer=="MRR") %>% pull(sig_to_init)
r_ic <- t_all %>% filter(cancer_type=="Lung",
                         Reviewer=="MRR") %>% pull(init_to_complete)
s_si <- t_all %>% filter(cancer_type=="Lung",
                         Reviewer=="SK") %>% pull(sig_to_init)
s_ic <- t_all %>% filter(cancer_type=="Lung",
                         Reviewer=="SK") %>% pull(init_to_complete)

t_all %>% 
  filter(cancer_type=="Lung") %>% 
  ggplot(aes(init_to_complete, factor(Reviewer))) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), color = "#777777") +
  geom_jitter(width = 0, height=0.25, alpha = 0.4) +
  scale_x_sqrt()

t_all %>% 
  filter(cancer_type=="Lung") %>% 
  ggplot(aes(sig_to_init, factor(Reviewer))) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), color = "#777777") +
  geom_jitter(width = 0, height=0.25, alpha = 0.4) +
  scale_x_sqrt()

wilcox.test(r_si, s_si)
wilcox.test(r_ic, s_ic)
```


